---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidync)
library(tidyverse)
library(R.utils)
library(lubridate)
library(purrr)
library(maps)
```

# PRECIPITATION GRAPH
``` {r}

#gunzip("cru_ts3.21.2001.2010.pre.dat.nc.gz", remove=FALSE)

file <- "cru_ts3.21.2001.2010.pre.dat.nc"
```

```{r}
lon  <- 
    tidync(file) %>% activate("D0") %>% hyper_tibble()
lat <-  
    tidync(file) %>% activate("D1") %>% hyper_tibble()
time <- 
    tidync(file) %>% activate("D2") %>% hyper_tibble()
```


```{r}
location_data <-
  tidync(file) %>% activate("pre") %>%
  hyper_filter(lon = lon == 149.25, lat = lat== -30.75) %>%
  hyper_tibble()
```


```{r}
location_data %>%
   ggplot(aes(x=time, y=pre)) + 
    geom_point() 
```


```{r}
location_data %>%
    mutate(pre_mean=mean(pre)) %>% 
   ggplot(aes(x=time, y=pre)) +
    geom_point() +
    geom_hline(aes(yintercept = pre_mean), color="red")
    

location_data %>%
    mutate(pre_mean=mean(pre)) %>% 
    ggplot(aes(x=time, y=pre)) + 
    geom_line() +
    geom_hline(aes(yintercept = pre_mean), color="red")
```



# TEMPERATURE GRAPH
```{r}
if(!file.exists(file)){
gunzip("cru_ts3.21.2001.2010.tmp.dat.nc", remove=FALSE)
}

file.2 <- "cru_ts3.21.2001.2010.tmp.dat.nc"
```


```{r}
lon  <- 
    tidync(file.2) %>% activate("D0") %>% hyper_tibble()
lat <-  
    tidync(file.2) %>% activate("D1") %>% hyper_tibble()
time <- 
    tidync(file.2) %>% activate("D2") %>% hyper_tibble()
```

```{r}
location_data <-
  tidync(file.2) %>% activate("tmp") %>%
  hyper_filter(lat = abs(lat - -30.75) == min(abs(lat - -30.75))) %>%
  hyper_filter(lon = abs(lon - 149.25) == min(abs(lon - 149.25))) %>%
  hyper_tibble()%>%
  mutate(time = as.Date(time, origin = "1900-01-01"))

```

```{r}
location_data %>%
  mutate(mean_temp = mean(tmp)) %>%
  ggplot(aes(x = time, y = tmp)) +
  geom_line() +
  geom_hline(aes(yintercept = mean_temp), linetype = "dashed", size = 2 ) +
  theme_classic()
```

```{r}
#y ~ 1 + c
#Y ~ 1
#y ~ month
lm(tmp ~lat, location_data)
lm1 <- lm(tmp ~1, location_data)
```


```{r}
coef(lm1)
location_data %>%
  mutate(mean_temp = mean(tmp), std_err_temp = coef(summary(lm1))[, "Std. Error"]) %>%
  mutate(estimated_mean = coef(lm1), lower_ci = estimated_mean - 1.96*std_err_temp, higher_ci = estimated_mean + 1.96*std_err_temp,
         low.quantile = estimated_mean - 1.96*sd(tmp),
         up.quantile = estimated_mean + 1.96*sd(tmp)) %>%
  ggplot(aes(x = time, y = tmp)) +
  geom_line() +
  geom_hline(aes(yintercept = mean_temp), linetype = "dashed", size = 2 ) +
  geom_hline(aes(yintercept = estimated_mean), linetype = "dotted", col = "red", size = 2 ) +
  geom_ribbon(aes(ymin = low.quantile, ymax = up.quantile), alpha = 0.1, fill = "yellow") +
  geom_ribbon(aes(ymin = lower_ci, ymax = higher_ci), alpha = 0.1, fill = "red") +
  theme_classic()
summary(lm1)$adj.r.squared
```


```{r}
location_data_month <-
location_data %>%
    mutate(date = ymd(time)) %>% 
  mutate_at(vars(date), funs(year, month, day)) %>%
    mutate(month= month(month, label= TRUE), month_character= as.character(month))
```


```{r}
fit <- lm(tmp ~ month_character, location_data_month)
summary(fit)
```

```{r}
predict(fit)
```

```{r}
location_data_predicted <-
location_data_month %>%
    mutate(predicted_temp = predict(fit))
```

```{r}
ggplot(location_data_predicted, aes(y = tmp, x = time)) +
    geom_line() + 
    geom_line(aes(y = predicted_temp), colour = 'red')
```
```{r}
location_data <-
  tidync(file.2) %>% activate("tmp") %>%
  hyper_tibble()%>%
  mutate(time = as.Date(time, origin = "1900-01-01")) %>%
  mutate(month= month(time, label= TRUE))

```

```{r}
calc_r_squared <- function(data, target_lat, target_lon) {
    
    data_slice<- data %>%
        filter(lon== target_lon & lat== target_lat)
    lm_data_slice<-lm(tmp~month, data_slice)
    summary(lm_data_slice)$adj.r.squared
}

calc_r_squared(data=location_data, target_lat=-54.75, target_lon=-68.25)
```


```{r}
calc_predictions <- function(data, target_lat, target_lon) {
    data_slice<- data %>%
        filter(lon== target_lon & lat== target_lat)
    lm_fit <- lm(tmp ~ month, data_slice)
    predict(lm_fit)
}
calc_predictions(data=location_data, target_lat=-54.75, target_lon=-71.75)
```
```{r}
location_data <-
  tidync(file.2) %>% activate("tmp") %>%
  hyper_tibble()%>%
  mutate(time = as.Date(time, origin = "1900-01-01")) %>%
  mutate(month= month(time, label= TRUE))
```

```{r}
# map functions using purrr

library(purrr)

# add1 <- function(x) {x+1}

# x <- (1:100)

# ?map

# map(x, add1) %>% str()

# map_dbl(x, add1) %>% str()
```
# Original Function
```{r}
calc_r_squared <- function(data, target_lat, target_lon) {

 

    data_slice<- data %>%

 

        filter(lon== target_lon & lat== target_lat)

 

    lm_data_slice<-lm(tmp~month, data_slice)

 

    summary(lm_data_slice)$adj.r.squared}

 

calc_r_squared(data =location_data, target_lat= 46.25, target_lon= 2.25)
```

 

```{r}
?nest

location_data %>% nest(tmp_data=c(tmp, time, month)) -> location_data_nest

location_data_nest$tmp_data[[3]]

calc_r_squared_map <- function(data) {

 

    tmp_data_fit<-lm(tmp~month, data)

 

    summary(tmp_data_fit)$adj.r.squared
    }
```

 

# Dont need to define arguments, because function takes a single argument
```{r}
location_data_nest <- location_data_nest %>% mutate(r_squared= map_dbl(tmp_data, calc_r_squared_map))
```
# creating a model, storing fitted models in 1 column, and r2 in another. Can be replicated for other variables
```{r}
fit_tmp_model <- function(data) {

 

   lm(tmp~month, data)
}
extract_r2 <- function(fit){
    summary(fit)$adj.r.squared
    
}
location_data_nest_r2 <- location_data_nest_sliced 
   %>% slice(1:100000) %>%
    mutate(
        fit_tmp= map(tmp_data_fit, fit_tmp_model), 
        tmp_r_squared= map_dbl(fit_tmp, extract_r2))
```





```{r}
TOTAL_PRECIP <- "cru_ts3.22.1901.2013.pre.dat.nc"
TOTAL_TMP <- "cru_ts3.22.1901.2013.tmp.dat.nc"
```
 
 
```{r}
lon  <-
    tidync(TOTAL_PRECIP) %>% activate("D0") %>% hyper_tibble()
lat <- 
    tidync(TOTAL_PRECIP) %>% activate("D1") %>% hyper_tibble()
time <-
    tidync(TOTAL_PRECIP) %>% activate("D2") %>% hyper_tibble() %>%
    mutate(time_ymd = as.Date(time, origin = "1900-01-01")) %>%
    mutate(month= month(time_ymd, label= TRUE)) %>%
    mutate(year= year(time_ymd))
```
 
```{r}
Total_data_precip <-
  tidync(TOTAL_PRECIP) %>% activate("pre") %>%
    hyper_filter(time = time >= 29234, lat = abs(lat) < 70) %>%
  hyper_tibble() %>%
  mutate(time = as.Date(time, origin = "1900-01-01")) %>%
    mutate(month= month(time, label= TRUE))%>%
    mutate(year= year(time))
```

```

```{r}
#Creating function using full data set
TOTAL_PRECIP <- "cru_ts3.22.1901.2013.pre.dat.nc"
TOTAL_TMP <- "cru_ts3.22.1901.2013.tmp.dat.nc"
```
 
 
```{r}
lon <-
    tidync(TOTAL_PRECIP) %>% activate("D0") %>% hyper_tibble()
lat <- 
    tidync(TOTAL_PRECIP) %>% activate("D1") %>% hyper_tibble()
time <-
    tidync(TOTAL_PRECIP) %>% activate("D2") %>% hyper_tibble() %>%
    mutate(time_ymd = as.Date(time, origin = "1900-01-01")) %>%
    mutate(month= month(time_ymd, label= TRUE)) %>%
   mutate(year= year(time_ymd))
    
```
 
 
```{r}
Total_data_precip <-
  tidync(TOTAL_PRECIP) %>% activate("pre") %>%
    hyper_filter(time = time >= 29234, lat = abs(lat) < 70) %>%
  hyper_tibble() %>%
  mutate(time = as.Date(time, origin = "1900-01-01")) %>%
    mutate(month= month(time, label= TRUE))%>%
    mutate(year= year(time))
```
 
```{r}
#Nesting precipitation data
Total_data_precip %>% nest(pre_data=c(pre,time,month,year)) -> precip_data_nested
```
 
```{r}
precip_data_nested$pre_data[[4]]
```
 
 
```{r}
fit_total_pre_model <- function(data) {
   
      lm(pre~month, data)
}
pre_extract_r2 <- function(fit) {
   
    summary(fit)$adj.r.squared
}
   
precip_data_nested <- precip_data_nested %>%
    slice(1:3000) %>% mutate(
        fit_pre = map(pre_data,fit_total_pre_model),
        pre_r_squared = map_dbl(fit_pre, extract_r2))
 
```
 
 
#Create plot for precip
```{r}
ggplot(precip_data_nested, aes(y = lat, x = lon))+
    geom_point()
```
```

